---
title: "R Notebook"
output: html_notebook
---

# Association between diabetes and neurodegeneration

## Does diabetes affect the time it takes to develop neurodegenerative diseases?

In order to answer this question, we need to find the patients answered yes in the question if they ever have diabetes. 

```{r}
library(dplyr)
library(ggplot2)
readDatFile <- function(path) {
  read.csv(path, stringsAsFactors = F, header = T, sep='\t')
}
WHI.f134 <- readDatFile('data/WHI/Medical History/f134_ctos_inv/f134_ctos_inv.dat')
```
```{r}
print(sum(WHI.f134$F134DIAB, na.rm=T))
print(sum(WHI.f134$F134PARKINS, na.rm=T))
```

There are in total 13840 participants who answered yes. There is also a Parkisons disease column, and there are 1260 individuals answered yes.

Now take a look at all the individuals in the observation study follow up survey, and collect the neurodegeneration and macular degeneration variables

```{r}
WHI.baseline <- readDatFile('data/WHI/Medical History/f30_ctos_inv/f30_ctos_inv.dat')
WHI.os3 <- readDatFile('data/WHI/OS Followup/f143_av3_os_inv/f143_av3_os_inv.dat')
WHI.os4 <- readDatFile('data/WHI/OS Followup/f144_av4_os_inv/f144_av4_os_inv.dat')
WHI.os5 <- readDatFile('data/WHI/OS Followup/f145_av5_os_inv/f145_av5_os_inv.dat')
WHI.os6 <- readDatFile('data/WHI/OS Followup/f146_av6_os_inv/f146_av6_os_inv.dat')
WHI.os7 <- readDatFile('data/WHI/OS Followup/f147_av7_os_inv/f147_av7_os_inv.dat')
WHI.baseline2 <- readDatFile('data/WHI/Demographics/f2_ctos_inv/f2_ctos_inv.dat')
```


Now join the followup data with selected column with the main WHI f134 data (left join).

```{r}
WHI.merged <- WHI.baseline2[, c('ID', 'AGE', 'F2DAYS', 'DIAB', 'DIABAGE', 'DIABCOMA', 'DBDIETF2', 'INSULIN', 'INSULINW', 'DIABPILL', 'DIABNW')] %>%
  filter(!is.na(DIAB)) %>%
  left_join(WHI.f134[, c('ID', 'F134DAYS', 'F134PARKINS', 'F134DIAB')]) %>%
  left_join(WHI.baseline[,c('ID', 'F30DAYS', 'ALZHEIM', 'MS', 'PARKINS', 'ALS', 'GLAUCOMA')]) %>%
  left_join(WHI.os3[,c('ID', 'F143DAYS', 'MACDEGNR_3', 'ALZHEIM_3', 'MS_3', 'PARKINS_3', 'ALS_3')]) %>%
  left_join(WHI.os4[,c('ID', 'F144DAYS', 'MACDEGNR_4', 'ALZHEIM_4', 'MS_4', 'PARKINS_4', 'ALS_4')]) %>%
  left_join(WHI.os5[,c('ID', 'F145DAYS', 'MACDEGNR_5', 'ALZHEIM_5', 'MS_5', 'PARKINS_5', 'ALS_5')]) %>%
  left_join(WHI.os6[,c('ID', 'F146DAYS', 'MACDEGNR_6', 'ALZHEIM_6', 'MS_6', 'PARKINS_6', 'ALS_6')]) %>%
  left_join(WHI.os7[,c('ID', 'F147DAYS', 'MACDEGNR_7', 'ALZHEIM_7', 'MS_7', 'PARKINS_7', 'ALS_7')]) 
```


Now remove those that completely have no followup.

```{r}
nodata.idx <- apply(WHI.merged, 1, function(x) {
  sum(is.na(x[seq(20, 49)])) == 30
})
WHI.filtered <- WHI.merged[!nodata.idx,]
```

After removing those that have absolutely no followup, only half of the original population is left. There are 62590 people in the dataset.

## Estimate the time to parkison's disease using mixed model

Use `nlme` for nonlinear mixed effect model

```{r}
library(lme4)

```



